Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-04-11 paper/paper.tex
update

\cyan{a comment} and \xcyan{a resolved comment}

@@

The source code in Julia language for this paper can be found in our Github repository~\cite{GraphTensorNetworks}. 
There is also a short introduction to this repository as well a gist to show how it works in the Appendix~\ref{sec:technical}.
We expect our tool can be used to understand and study many interesting applications of independent sets and beyond.
We also hope the toolkit we built, including tensor network contraction order optimization and efficient tropical matrix multiplication, can be helpful to other scientific software.

@@

This appendix covers some technical guides for productivity, including an introduction to an open source package \href{https://github.com/Happy-Diode/GraphTensorNetworks.jl}{GraphTensorNetworks}~\cite{GraphTensorNetworks} implementing the algorithms in this paper and a gist about how this package is implemented.
One can install \texttt{GraphTensorNetworks} a Julia REPL, by first type \colorbox{lightgray}{\texttt{]}} to enter the \texttt{pkg>} mode and then type
\begin{lstlisting}
pkg> add GraphTensorNetworks
\end{lstlisting}
followed by an \colorbox{lightgray}{\texttt{<ENTER>}} key.
To use it for solving solution space properties, just go back to the normal mode (type \colorbox{lightgray}{\texttt{<BACKSPACE>}}) and type

\begin{lstlisting}
julia> using GraphTensorNetworks, Graphs

@@

Here the main function \texttt{solve} takes three inputs, the problem instance of type \texttt{IndependentSet}, the property instance of type \texttt{GraphPolynomial} and an optional key word argument \texttt{usecuda} to decide use GPU or not.
If one wants to use GPU to accelerate the computation, ``\texttt{using CUDA}'' must uncommented.
The problem instance takes four arguments to initialize, the only positional argument is the graph instance one wants to solve, the key word argument \texttt{optimizer} is for specifying the tensor network optimization algorithm, the key word argument \texttt{weights} is for specifying the weights of vertices as either a vector or \texttt{NoWeight()} and the keyword argument \texttt{openvertices} is for specifying the degrees of freedom not summed over.
Here, we use \texttt{TreeSA} method as the tensor network optimizer, and leave \texttt{weights} and \texttt{openvertices} the default values. The \texttt{TreeSA} algorithm, which was invented in Ref.~\cite{Kalachev2021}, performs the best in most of our applications.
The first execution of this function will be a bit slow due to Julia's just in time compiling.
After that, the subsequent runs will be fast.
The following diagram lists possible combinations of input arguments, where functions in the \texttt{Graph} are mainly defined in package \texttt{Graphs}, and the rest can be found in \texttt{GraphTensorNetworks}.

@@

They can be installed in a similar way to \texttt{GraphTensorNetworks}.
After installing the required packages, one can open a Julia REPL, copy and paste the following code snippet into it.

>>>>

@@

The source code in Julia language for this paper can be found in our Github repository~\cite{GenericTensorNetworks}. 
There is also a short introduction to this repository as well a gist to show how it works in the Appendix~\ref{sec:technical}.
We expect our tool can be used to understand and study many interesting applications of independent sets and beyond.
We also hope the toolkit we built, including tensor network contraction order optimization and efficient tropical matrix multiplication, can be helpful to other scientific software.

@@

This appendix covers some technical guides for productivity, including an introduction to an open source package \href{https://github.com/Happy-Diode/GenericTensorNetworks.jl}{GenericTensorNetworks}~\cite{GenericTensorNetworks} implementing the algorithms in this paper and a gist about how this package is implemented.
One can install \texttt{GenericTensorNetworks} a Julia REPL, by first type \colorbox{lightgray}{\texttt{]}} to enter the \texttt{pkg>} mode and then type
\begin{lstlisting}
pkg> add GenericTensorNetworks
\end{lstlisting}
followed by an \colorbox{lightgray}{\texttt{<ENTER>}} key.
To use it for solving solution space properties, just go back to the normal mode (type \colorbox{lightgray}{\texttt{<BACKSPACE>}}) and type

\begin{lstlisting}
julia> using GenericTensorNetworks, Graphs

@@

Here the main function \texttt{solve} takes three inputs, the problem instance of type \texttt{IndependentSet}, the property instance of type \texttt{GraphPolynomial} and an optional key word argument \texttt{usecuda} to decide use GPU or not.
If one wants to use GPU to accelerate the computation, ``\texttt{using CUDA}'' must uncommented.
The problem instance takes four arguments to initialize, the only positional argument is the graph instance one wants to solve, the key word argument \texttt{optimizer} is for specifying the tensor network optimization algorithm, the key word argument \texttt{weights} is for specifying the weights of vertices as either a vector or \texttt{NoWeight()} and the keyword argument \texttt{openvertices} is for specifying the degrees of freedom not summed over.
Here, we use \texttt{TreeSA} method as the tensor network optimizer, and leave \texttt{weights} and \texttt{openvertices} the default values. The \texttt{TreeSA} algorithm, which was invented in Ref.~\cite{Kalachev2021}, performs the best in most of our applications.
The first execution of this function will be a bit slow due to Julia's just in time compiling.
After that, the subsequent runs will be fast.
The following diagram lists possible combinations of input arguments, where functions in the \texttt{Graph} are mainly defined in package \texttt{Graphs}, and the rest can be found in \texttt{GenericTensorNetworks}.

@@

They can be installed in a similar way to \texttt{GenericTensorNetworks}.
After installing the required packages, one can open a Julia REPL, copy and paste the following code snippet into it.

### 2022-04-11 paper/paper_before_maddie.tex
update

\section{Discussion and conclusion}
In this paper, we introduce a method to use generic programming and tensor networks to compute a number of different solution space properties of the maximum independent set problem,
including the largest several independent sets sizes, the number of independent sets of a given size, and
the enumeration of independent sets of a given size.
For each solution space property, we design a commutative semiring algebra for the tensor element type.
With different algebras, the contraction of the same tensor network gives us the different properties.
The data types introduced in the main text is summarized in the diagram in \Fig{fig:venn-diagram}.
Other than the case with real numbers, we have polynomials and truncated polynomials.
We combine them with the bit string algebra, the set algebra, and expression tree algebra for finding, enumeration, and sampling of independent sets at a given size.
These algebras are of generic use: they can be utilized to compute properties of maximal independent sets and a variety of other combinatorial problems such as the matching problem, the k-coloring problem, the max-cut problem, and the set packing problem, as detailed in \App{sec:otherproblems}. 
Moreover, since the independence polynomial is closely related to the matching polynomial~\cite{Levit2005}, the clique polynomial~\cite{Hoede1994}, and the vertex cover polynomial~\cite{Akbari2013},
our algorithm to compute the independence polynomial can also be used to compute these graph polynomials.
By adapting to the style of generic programming, we can implement all the algorithms without much effort.
We show some of the Julia language implementations in Appendix~\ref{sec:technical}.
A complete implementation can be found in our Github repository~\cite{GraphTensorNetworks}. 
We expect our tool can be used to understand and study many interesting applications of independent sets and beyond. We also hope this approach of generic tensor networks can inspire future works on tensor network computations.

@@

For better performance, we recommend checking our GitHub repository for the full featured version:
\href{https://github.com/Happy-Diode/GraphTensorNetworks.jl}{https://github.com/Happy-Diode/GraphTensorNetworks.jl}.
It can be installed in a similar style to other packages.

>>>>

\section{Discussion and conclusion}
In this paper, we introduce a method to use generic programming and tensor networks to compute a number of different solution space properties of the maximum independent set problem,
including the largest several independent sets sizes, the number of independent sets of a given size, and
the enumeration of independent sets of a given size.
For each solution space property, we design a commutative semiring algebra for the tensor element type.
With different algebras, the contraction of the same tensor network gives us the different properties.
The data types introduced in the main text is summarized in the diagram in \Fig{fig:venn-diagram}.
Other than the case with real numbers, we have polynomials and truncated polynomials.
We combine them with the bit string algebra, the set algebra, and expression tree algebra for finding, enumeration, and sampling of independent sets at a given size.
These algebras are of generic use: they can be utilized to compute properties of maximal independent sets and a variety of other combinatorial problems such as the matching problem, the k-coloring problem, the max-cut problem, and the set packing problem, as detailed in \App{sec:otherproblems}. 
Moreover, since the independence polynomial is closely related to the matching polynomial~\cite{Levit2005}, the clique polynomial~\cite{Hoede1994}, and the vertex cover polynomial~\cite{Akbari2013},
our algorithm to compute the independence polynomial can also be used to compute these graph polynomials.
By adapting to the style of generic programming, we can implement all the algorithms without much effort.
We show some of the Julia language implementations in Appendix~\ref{sec:technical}.
A complete implementation can be found in our Github repository~\cite{GenericTensorNetworks}. 
We expect our tool can be used to understand and study many interesting applications of independent sets and beyond. We also hope this approach of generic tensor networks can inspire future works on tensor network computations.

@@

For better performance, we recommend checking our GitHub repository for the full featured version:
\href{https://github.com/Happy-Diode/GenericTensorNetworks.jl}{https://github.com/Happy-Diode/GenericTensorNetworks.jl}.
It can be installed in a similar style to other packages.
